
# Reproducible Research: Peer Assessment 1 
# Activity Monitoring on Personal Movement
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
    
## Loading and preprocessing the data

## 1.Load the data
```{r,echo=TRUE}
actv <- read.csv("../RepData_PeerAssessment1-master/activity.csv", header = T, sep = ",")
str(actv)
```

## 2.Process/transform the data (if necessary) into a 
## format suitable for your analysis
```{r,echo=TRUE}
actv[,'date'] <- as.Date(actv[,'date'])
str(actv)
```


## What is mean total number of steps taken per day?
#### For this part of the assignment, you can ignore the missing values in the dataset.

## 1.Calculate the total number of steps taken per day
## 2.Make a histogram of the total number of steps taken each day. 
#### note the difference between barplot
```{r,echo=TRUE}
totStep <- aggregate(actv[c("steps")], 
                         list(date = actv$date), sum ,na.rm=T)

hist(totStep$steps,xlab="total steps",ylab="Freqency",main="total steps per day",
     col="blue" )
```

## 3.Calculate and report the mean and median of the total number of steps taken per day
```{r,echo=TRUE}
tmp <- totStep[,2]
meanStep <- round(mean(tmp,na.rm=TRUE))
print(c("the mean of total number of steps is",meanStep))

medianStep <- round(median(tmp,na.rm=TRUE))
print(c("the median of total number of steps is",medianStep))
```

## What is the average daily activity pattern?

## 1.Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 
## the average number of steps taken, averaged across all days (y-axis)
```{r,echo=TRUE}
avgStepInterval <- aggregate(actv[c("steps")], list(interval=actv$interval),FUN=mean,na.rm=T)
plot(avgStepInterval,xlab="5-minute interval",ylab="average steps", type="l")
```

## 2.Which 5-minute interval,on average across all the days in the dataset,contains the maximum number of steps?
```{r,echo=TRUE}
maxinterval <- subset(avgStepInterval, 
                      avgStepInterval$steps==max(avgStepInterval$steps, na.rm = TRUE))
maxinterval
```

## Imputing missing values
#### Note that there are a number of days/intervals where there are missing values (coded as NA). 
#### The presence of missing days may introduce bias into some calculations or summaries of the data.

## 1. Calculate and report the total number of missing values in the dataset (i.e.the total number of rows withNAs)
```{r,echo=TRUE}
rowMiss <- sum(is.na(actv$steps))
print(c("rows with missing steps data is",rowMiss))
```

## 2. Devise a strategy for filling in all of the missing values in the dataset.
## The strategy does not need to be sophisticated.
## For example, you could use the mean/median 
## for that day, or the mean for that 5-minute interval, etc.
```{r,echo=TRUE}
forMiss <- aggregate(actv[c("steps")], list(interval=actv$interval), mean, na.rm=T)
colnames(forMiss)[colnames(forMiss)=="steps"] <- "meanv"

actv1 <- merge(actv,forMiss,by="interval",sort=TRUE)
colnames(actv1)[colnames(actv1)=='steps'] <- "osteps"
```


## 3. Create a new dataset that is equal to the original dataset but 
## with the missing data filled in.
```{r,echo=TRUE}
actv1[,"steps"] <- ifelse(is.na(actv1[,"osteps"]), actv1[,"meanv"], actv1[,"osteps"])
actv1$meanv = NULL
actv1$osteps = NULL
rowMiss <- sum(is.na(actv1$steps))
print(c("after adjustment rows of missing data",rowMiss))
```

## 4. Make a histogram of the total number of steps taken each day and 
## Calculate and report the mean and median total number of steps taken per day. 
## Do these values differ from the estimates from the first part of the assignment? 
## What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r,echo=TRUE}
totStep1 <- aggregate(actv1[c("steps")], 
                         list(date = actv$date), sum ,na.rm=T)

par(mfrow=c(1,2))
hist(totStep$steps,xlab="total steps each day",ylab="Freqency",
    main="with missing value",
     col="blue"  )

hist(totStep1$steps,xlab="total steps each day",ylab="Freqency",
     main="after adjust missing value",
     col="blue" )
```



```{r,echo=TRUE}
tmp <- totStep1[,2]
meanStep <- round(mean(tmp,na.rm=TRUE))
print(c("after adjustment, mean of total number of steps is",meanStep))

medianStep <- round(median(tmp,na.rm=TRUE))
print(c("after adjustment, median of total number of steps is",medianStep))


print(c("the mean has increase to be close to median"))
print(c("the impute method result in shift in the frequency distribution of total steps"))

```

## Are there differences in activity patterns between weekdays and weekends?
#### For this part the weekdays() function may be of some help here. 
#### Use the dataset with the filled-in missing values for this part.

## 1. Create a new factor variable in the dataset with two levels – 
## “weekday” and “weekend” indicating whether a given date is 
## a weekday or weekend day.
```{r,echo=TRUE}
actv1[,"weekd"] <- weekdays(actv1[,"date"])
actv1[,"weekdg"] <- ifelse((actv1[,"weekd"] %in% c("Sunday","Saturday")), "weekend", "weekday")
actv1[,"weekdg"] <-factor(actv1[,"weekdg"])
```

## 2. Make a panel plot containing a time series plot (i.e. type = "l") 
## of the 5-minute interval (x-axis) and the average number of steps taken, 
## averaged across all weekday days or weekend days (y-axis). 
## See the README file in the GitHub repository to see an example of 
## what this plot should look like using simulated data.

```{r,echo=TRUE}
 avgStepInterval1 <- aggregate(actv1[c("steps")], 
                             list(interval = actv1$interval, daytype = actv1$weekdg),
                             FUN=mean, na.rm=T)

library(lattice)
xyplot(steps ~ interval | daytype, data=avgStepInterval1, layout=c(1,2), 
       type="l", ylab = "average steps")
```

```{r,echo=TRUE}
print(c("except for 1 interval, the general trend of steps for weekend and weekday is the same"))
```
